### 2024-04-18 paper/paper.qmd
updating model

---
title: "How South Koreans View Democracy?"
subtitle: "An Analysis of Survey Data in 2016"
author: 
  - Hailey Jang 
thanks: "Code and data are available at: https://github.com/Hailey-Jang/How_South_Koreans_View_Democracy.git."
date: today
date-format: long
abstract: "This study explores the impact of globalization on democratic values in South Korea by employing a linear regression model to analyze responses from a nationwide questionnaire on citizens' perceptions. Through the quantitative analysis, the research uncovers a complex relationship between diverse factors and democratic perception, demonstrating how political dissatisfaction can diminish feelings of representation and reduce democratic participation. These findings underscore the importance of addressing economic and political dissatisfaction to foster a more inclusive and participative democratic environment in South Korea, contributing to our broader understanding of how globalization shapes political systems worldwide."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(knitr)
library(kableExtra)
```

# Introduction

In South Korea, a nation where I was born and raised, the rapid and profound impacts of globalization have redefined the socio-economic and political landscape. As democracy inherently upholds the principle of political equality, ensuring that every citizen has a voice in the political discourse, these global forces present significant challenges. South Korea, characterized by swift economic growth and transformative political shifts, provides a critical context for exploring these challenges. These changes impact more than just the economy; they deeply influence the democratic ethos, raising crucial questions about the sustainability of political equality amidst pervasive transformations. Thus, the citizens of South Korea find themselves navigating an altered political and economic terrain, necessitating a critical reevaluation of democratic values in response to these global dynamics.

This paper delves into the significant question of how economic and political dissatisfaction influence the democratic perceptions of South Korean citizens. By examining how global forces reflect in the populace's political engagement and representational sentiments, this study seeks to enrich our understanding of the contemporary democratic ethos in South Korea. To address these inquiries, this research utilizes data derived from a comprehensive questionnaire distributed to Korean citizens, capturing their perceptions of economic and political aspects. Employing linear regression models, the analysis focuses on identifying and interpreting the relationships between citizens' economic conditions, political contexts, and their democratic engagement. This methodological approach allows for a detailed exploration of how socio-economic factors influence perceptions and behaviours related to democracy in South Korea.

The findings of this study are vital for understanding the current state and future trajectory of democracy in regions undergoing rapid globalization. By highlighting how socio-economic dissatisfaction influence democratic engagement, this research offers insights that are crucial for policymakers, scholars, and citizens concerned with fostering a robust democratic society in the face of global changes.

This paper is structured to first provide a detailed literature review that sets the theoretical framework for understanding the impact of globalization on democratic values. Following this, @sec-data describes the survey design and analytical techniques employed. @sec-result present the analysis results, discuss the findings in the context of existing literature in @sec-discussion, and conclude with implications for policy and suggestions for future research. By offering a comprehensive exploration of how global trends and socio-economic transformations influence democratic perceptions and participation in South Korea, this study contributes valuable insights into the evolving nature of democracy in a globalized world.

# Data {#sec-data}
This project harnessed the capabilities of the statistical software, `R` [@citeR], throughout its execution. The `tidyverse` [@tidy] package was indispensable for data cleaning and manipulation tasks. The creation of figures and other data visualizations was adeptly handled using `ggplot2` [@ggplot2R], while `knitr` [@knitrR] and `kableExtra` [@kableextra] were utilized for their excellent data presentation features. File path management was conducted with `readr` [@readrR], and `rstanarm` [@rstanarm] was employed for statistical modeling, along with the `palmerpenguins` [@palmerpenguins] for illustrative purposes.

The data used in this study comes from the "Voter Political Consciousness Survey" conducted in South Korea following the 20th presidential election in 2016, under the auspices of The Comparative Study of Electoral Systems (CSES), a collaborative research project among national election studies worldwide. In this questionnaire, voters disclosed their economic and political perceptions. The raw dataset was large and unwieldy, so it was manually cleaned and renamed for readability. Variables were renamed as follows: E1006_NAM for the country name, E2002 for gender, E2001_Y for birth year, E3023 for satisfaction with democracy, E3007 for corruption, E3009 for political perception, and E3011 for economic perception. Unnecessary rows were deleted from the dataset.

## Data Summary

As show in @tbl-variables, the dependent variable will be based on the question, "On the whole, are you very satisfied, fairly satisfied, not very satisfied, or not at all satisfied with the way democracy works in South Korea?" The responses are coded on a four-point scale: 1. very satisfied, 2. fairly satisfied, 3. not very satisfied, and 4. not at all satisfied. For the independent variables, key variables on economic satisfaction will include responses to the question, "Would you say that over the past twelve months, the state of the economy in South Korea has gotten much better, gotten somewhat better, stayed about the same, gotten somewhat worse, or gotten much worse?" Answers are coded as 1. Gotten better, 2. stayed the same, 3. gotten worse, 4. don’t know. Additionally, perceptions of corruption will be assessed with the question, "Most politicians are trustworthy." Responses are coded as 1. Strongly agree, 2. Somewhat agree, 3. Neither, 4. Somewhat disagree, 5. Strongly disagree. Lastly, data on political changes will be obtained from the question regarding the performance of President Geun-hye Park's government in 2016: "Now thinking about the performance of the government of Geun-hye Park president in general, how good or bad a job do you think the government of Geun-hye Park president did over the past three years? Has she done a very good job?" Political satisfaction responses are coded as 1. Very good job, 2. Good job, 3. Bad job, 4. Very bad job.

```{r}
#| label: tbl-variables
#| tbl-cap: Summary of Variables
#| echo: false
#| warning: false

library(knitr)
library(kableExtra)

# Create the data frame with the variable summary
variable_summary <- data.frame(
  Types = c("The dependent variable", "The independent variables", "", ""),
  Variables = c("Dissatisfaction with democracy", "Dissatisfaction with economy", "Distrust level agianst politicians", "Dissatisfaction with politics"),
  Criterion = c(
    "1. Very satisfied, 2. Fairly satisfied, 3. Not very satisfied, 4. Not at all satisfied",
    "1. Gotten better, 2. Stayed the same, 3. Gotten worse, 4. Don’t know",
    "1. Strongly agree, 2. Somewhat agree, 3. Neither, 4. Somewhat disagree, 5. Strongly disagree",
    "1. Very good job, 2. Good job, 3. Bad job, 4. Very bad job"
  )
)

kable(variable_summary, booktabs = TRUE, escape = FALSE) %>%
  kable_styling(bootstrap_options = c("condensed"), full_width = F, position = "center") %>%
  column_spec(1, width = "2cm") %>%  
  column_spec(2, width = "5cm") %>% 
  column_spec(3, width = "7cm")
```

## Data Analysis
To further explore the variables, summary statistics will be presented to describe the relationships among them. @fig-gender-perceptions will show how men and women in South Korea perceive the state of the economy, corruption, and political state. @fig-age-perceptions will show how different generational layers in South Korea perceive the state of the economy, corruption, and political state.

In the @fig-gender-perceptions , we have a set of three bar graphs, each representing a different type of distrust level agianst Korean politicians, dissatisfaction of economy and politics. The bars are divided by gender, with one colour representing male respondents (labelled as '1') and another color representing female respondents (labelled as '2'). The y-axis shows the average perception score, which likely corresponds to a survey response scale. Each bar's height indicates the average score for that perception type among the respective gender group. Both genders have similar average scores for economic dissatisfaction, while there may be slight differences in distrust against politicians and political dissatisfaction.

```{r}
#| label: fig-gender-perceptions
#| fig-cap: Perceptions by Gender
#| echo: false
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readr)

# Load the data
analysis_data <- read_csv("/Users/hailey/Desktop/Democracy_Perception_in_South_Korea/data/analysis_data/analysis_data.csv")

analysis_data_long <- pivot_longer(analysis_data, 
                                   cols = c("Dissatisfaction_with_economy", "Distrust_level_agianst_politicians", "Dissatisfaction_with_politics"), 
                                   names_to = "Perception_Type", 
                                   values_to = "Perception_Score")

short_names <- c("Dissatisfaction_with_economy" = "Eco. Dissat.",
                 "Distrust_level_agianst_politicians" = "Pol. Distrust",
                 "Dissatisfaction_with_politics" = "Pol. Dissat.")

analysis_data_long$Perception_Type <- factor(analysis_data_long$Perception_Type,
                                             levels = names(short_names),
                                             labels = short_names[names(short_names)])

analysis_data_long$Gender <- factor(analysis_data_long$Gender, labels = c("Male", "Female"))

# Plotting Perceptions by Gender
ggplot(analysis_data_long, aes(x = factor(Gender), y = Perception_Score, fill = factor(Gender))) +
  geom_bar(stat = "summary", fun = mean, position = position_dodge()) +
  facet_wrap(~Perception_Type, scales = "free_y") +
  labs(x = "Gender", y = "Average Perception Score") +
  theme_minimal() +
  theme(axis.text.x = element_blank())
```

In the @fig-age-perceptions, the same three types of perceptions are displayed. This time, however, they are categorized by birth year, which are shown on the x-axis and range from 1931 to 1997. The darkness of the bars corresponds to age, with darker shades representing older age groups.The @fig-age-perceptions suggest a view of generational sentiments: while economic and political dissatisfaction remain relatively consistent across age groups, there is a notable divergence when it comes to distrust in politicians. The younger generations who exhibit higher levels of distrust indicates a growing skepticism about political figures among newer generations compared to their older counterparts. This shift could reflect the evolving political landscape and the impact of recent events on the trust levels of different age groups.

```{r}
#| label: fig-age-perceptions
#| fig-cap: Perceptions by Age Group
#| echo: false
#| warning: false

library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)

# Load the data
analysis_data <- read_csv("/Users/hailey/Desktop/Democracy_Perception_in_South_Korea/data/analysis_data/analysis_data.csv")

analysis_data_long <- pivot_longer(analysis_data, 
                                   cols = c("Dissatisfaction_with_economy", "Distrust_level_agianst_politicians", "Dissatisfaction_with_politics"), 
                                   names_to = "Perception_Type", 
                                   values_to = "Perception_Score")

short_names <- c("Dissatisfaction_with_economy" = "Eco. Dissat.",
                 "Distrust_level_agianst_politicians" = "Pol. Distrust",
                 "Dissatisfaction_with_politics" = "Pol. Dissat.")

analysis_data_long$Perception_Type <- factor(analysis_data_long$Perception_Type,
                                             levels = names(short_names),
                                             labels = short_names[names(short_names)])

ggplot(analysis_data_long, aes(x = Age, y = Perception_Score, fill = Age)) +
  geom_bar(stat = "summary", fun = mean, position = position_dodge()) +
  facet_wrap(~Perception_Type, scales = "free_y") +
  labs(x = "Age Group", y = "Average Perception Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 55, hjust = 1))
```

These figures collectively aim to shed light on the distribution of attitudes towards different aspects of democratic society among different demographic groups in South Korea. The variation in average scores across different genders and age groups may suggest underlying social or cultural factors that influence how different groups perceive economic conditions, the prevalence of corruption, and political performance.

# Model

## Model set-up

In the Bayesian linear regression model[@rstanarm], we attempt to predict $Satisfaction\_With\_Democracy$ (the dependent variable, denoted as $y_i$ for each observation $i$) using $Economic\_Perception$, $Corruption\_Perception$, and $Political\_Perception$ as independent variables. The $y_i$ values are assumed to be normally distributed with a mean of $\mu_i$ and a common standard deviation $\sigma$. The mean $\mu_i$ is a linear combination of the independent variables, represented by the equation $\mu_i = \beta_0 + \beta_1 \cdot Economic\_Perception_i + \beta_2 \cdot Corruption\_Perception_i + \beta_3 \cdot Political\_Perception_i$. Here, $\beta_0$ is the intercept, the expected value of $Satisfaction\_With\_Democracy$ when all independent variables are at zero, reflecting a baseline level of satisfaction without the influence of the predictors. The coefficients $\beta_1$, $\beta_2$, and $\beta_3$ are the slopes that measure the expected change in $Satisfaction\_With\_Democracy$ for a one-unit change in their respective independent variables, assuming other variables are held constant. These regression coefficients are assumed to follow normal distributions with means of zero and standard deviations of 2.5, indicating that before observing the data, we believe it is equally likely that the variables could have a positive or negative association with satisfaction, with a moderate level of certainty. The standard deviation $\sigma$ of the dependent variable's distribution captures the variability of $Satisfaction\_With\_Democracy$ around its mean that is not explained by the independent variables. An Exponential prior with a rate of 1 is assigned to $\sigma$, which suggests we expect smaller values of $\sigma$ (indicating less noise) to be more likely than larger ones. This setup of priors and the model framework allow us to incorporate prior beliefs and uncertainties about the parameters, which are then updated with the data to provide a posterior distribution reflecting our updated beliefs.

\begin{align*}
y_i &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 x_i \\
\beta_0 &\sim \text{Normal}(0, 2.5) \\
\beta_1 &\sim \text{Normal}(0, 2.5) \\
\sigma &\sim \text{Exponential}(1)
\end{align*}

### Model justification

# Result{#sec-result}

By examining the statistical analysis of @tbl-modelresults, we can explore the impact of economic and political satisfaction changes on the formation of citizens' perceptions. The results of the two models indicate that perceptions of corruption and economics do not significantly influence the formation of perceptions regarding the realization of democracy. However, it was observed that political satisfaction does have a meaningful effect on the formation of democratic realization perceptions. Considering that political perception has a positive relationship with the perception of democratic realization, it can be interpreted that people who perceive a high level of political satisfaction have a positive perception of the realization of democracy.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
second_model <-
  readRDS(file = here::here("models/second_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model,
    "Second model" = second_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion{#sec-discussion}

## Democracy Perception by Different Variables {#sec-first-point}
@fig-economic-perceptions is a scatter plot overlaid with a linear regression trend line, depicting the relationship between Economic dissatisfaction and dissatisfaction with Democracy.

The outcome of the graph suggests that there is a positive association between  dissatisfaction with economics and dissatisfaction with democracy, as indicated by the upward slope of the trend line. Points on the graph tend to shift from lighter to darker as they move right, suggesting that increased economic dissatisfaction correlates with a higher dissatisfaction with democracy, particularly as age increases. This could imply that older individuals, who potentially have more experience with economic fluctuations and political systems, might perceive a stronger link between economic and democratic dissatisfaction.

```{r}
#| label: fig-economic-perceptions
#| fig-cap: Democracy Perceptions by Economic Satisfaction
#| echo: false
#| warning: false

# Load the necessary libraries
library(ggplot2)
library(readr)

# Load the data
analysis_data <- read_csv("/Users/hailey/Desktop/Democracy_Perception_in_South_Korea/data/analysis_data/analysis_data.csv")

ggplot(analysis_data, aes(x = Dissatisfaction_with_economy, y = Dissatisfaction_with_democracy, color = Age)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = TRUE, color = "blue") +  
  scale_color_viridis_c(option = "D") +  
  labs(x = "Dissatisfaction with Economics", 
       y = "Dissatisfaction with Democracy", 
       title = "Democracy Perceptions by Economic Satisfaction") +
  theme_minimal() +
  theme(legend.position = "bottom")  
```

@fig-politicians-perceptions shows a correlation between mistrust in politicians and dissatisfaction with democracy, with a gradual upward trend line signifying that as mistrust against politicians increases, so does dissatisfaction with democracy. Notably, the younger age groups (lighter colors) seem to be concentrated at higher levels of mistrust, suggesting that younger individuals may be more inclined to distrust politicians and consequently feel more dissatisfied with democracy.

```{r}
#| label: fig-politicians-perceptions
#| fig-cap: Democracy Perceptions by Corruption Perception
#| echo: false
#| warning: false

ggplot(analysis_data, aes(x = Distrust_level_agianst_politicians, y = Dissatisfaction_with_democracy, color = Age)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = TRUE, color = "blue") +  
  scale_color_viridis_c(option = "D") +  
  labs(x = "Mistrust against Politicians", 
       y = "Dissatisfaction with Democracy") +
  theme_minimal() +
  theme(legend.position = "bottom")  
```

The @fig-political-perceptions displays a clear positive correlation between dissatisfaction with politics and dissatisfaction with democracy, indicated by the ascending trend line. As political dissatisfaction increases, so does dissatisfaction with democracy, suggesting that negative feelings towards the political system are closely linked with negative perceptions of democratic governance. The gradient color, representing age groups from older (darker) to younger (lighter), reveals a relatively even distribution across the spectrum of political dissatisfaction, signifying that these sentiments are pervasive across different ages without a visible age-related trend in the data. This indicates a general consensus across generations regarding the relationship between political dissatisfaction and democratic discontent.

```{r}
#| label: fig-political-perceptions
#| fig-cap: Democracy Perceptions by Political Perception
#| echo: false
#| warning: false

ggplot(analysis_data, aes(x = Dissatisfaction_with_politics, y = Dissatisfaction_with_democracy, color = Age)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = TRUE, color = "blue") +  
  scale_color_viridis_c(option = "D") +  
  labs(x = "Dissatisfaction with Politics", 
       y = "Dissatisfaction with Democracy") +
  theme_minimal() +
  theme(legend.position = "bottom")  
```
## Democracy Perception
The @fig-slop-comparison illustrates the estimated slopes of three different regression models, each representing the impact of a specific type of dissatisfaction—economic, political, and distrust against politicians—on dissatisfaction with democracy. The plot reveals that all three factors have a positive relationship with dissatisfaction with democracy, as indicated by the fact that all the estimated coefficients are above zero. Economic dissatisfaction has the smallest estimated impact, whereas political dissatisfaction has the highest, with distrust against politicians falling in between. The vertical lines represent the confidence intervals for each estimate, and none of them cross the horizontal line at zero, suggesting that all three relationships are statistically significant. The differences in the lengths of these lines indicate varying degrees of certainty about the estimates, with economic dissatisfaction showing the widest interval, implying the least precision.

```{r}
#| label: fig-slop-comparison
#| fig-cap: Comparison of Slopes from Different Regression Models
#| echo: false
#| warning: false

library(dplyr)
library(broom)

# Assuming analysis_data is already loaded and contains the variables
# Fit separate linear models for each factor
models <- list(
  econ = lm(Dissatisfaction_with_democracy ~ Dissatisfaction_with_economy, data = analysis_data),
  pol = lm(Dissatisfaction_with_democracy ~ Dissatisfaction_with_politics, data = analysis_data),
  trust = lm(Dissatisfaction_with_democracy ~ Distrust_level_agianst_politicians, data = analysis_data)
)

# Extract the coefficients with confidence intervals
model_coefficients <- bind_rows(
  lapply(models, function(model) {
    tidy(model, conf.int = TRUE) %>%
      filter(term != "(Intercept)") # Remove the intercept
  }),
  .id = "model"
)

# Rename the factor levels for better graph labeling
model_coefficients$model <- factor(model_coefficients$model, levels = c("econ", "pol", "trust"),
                                   labels = c("Economic Dissatisfaction", "Political Dissatisfaction", "Distrust Against Politicians"))

# Plot the coefficient for each model
ggplot(model_coefficients, aes(x = model, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Factor", y = "Estimated Slope (Coefficient)") +
  theme_minimal()
```

## Conclusions

The study presents an examination of how socio-economic and political dissatisfaction influence the democratic sentiments of South Koreans amidst the multifaceted challenges of globalization. By utilizing linear regression models to parse out the effects of these variables on democratic contentment, it delineates a clear pattern where political dissatisfaction holds the most sway in eroding democratic satisfaction, outpacing the influence of distrust in politicians and economic dissatisfaction. These findings illuminate the critical role of political harmony and economic stability in underpinning the health of a democracy. As South Korea faces the winds of global change, such insights are vital for crafting policies that shore up democratic integrity and public trust.

However, the research is not without limitations. It operates on the assumption that the survey responses accurately reflect the intricate reality of individual perceptions, which may not encapsulate the full spectrum of socio-political nuances. The methodology, while robust, cannot account for unmeasured variables that may confound the relationships explored. Furthermore, the cross-sectional nature of the survey data limits the ability to draw causal inferences. Despite these weaknesses, the study contributes significantly to our understanding of democracy in a rapidly evolving socio-economic landscape. It beckons a call for future research to adopt longitudinal approaches and consider a broader array of variables, thereby enriching the discourse on sustaining democratic values amidst the complexities of the modern world.

\newpage

# References